---
title: "HW 3"
author: "null"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warnings = FALSE, fig.align = 'center',  eval = TRUE)
```

We will be predicting whether the housing price is expensive or not using the `sahp` dataset in the **r02pro** package. 

You can run the following code to prepare the analysis.
```{r}
library(r02pro)     #INSTALL IF NECESSARY
library(tidyverse)  #INSTALL IF NECESSARY
library(MASS)
my_sahp <- sahp %>% 
  na.omit() %>%
  mutate(expensive = sale_price > median(sale_price)) %>%
  dplyr::select(gar_car, liv_area, oa_qual, expensive) 
my_sahp_train <- my_sahp[1:100, ]
my_sahp_test <- my_sahp[-(1:100), ]
```

Please answer the following questions.

#### Q1
    
a. Using the training data `my_sahp_train` to fit a logistic regression model of `expensive` on each variable (`gar_car`, `liv_area`, `oa_qual`) separately. For each logistic regression, compute the training and test error. Which variable leads to the smallest training error? Which variable leads to the smallest test error?

    **Solution:**

b. Using the training data `my_sahp_train` to fit a logistic regression model of `expensive` on all three variables (`gar_car`, `liv_area`, `oa_qual`). Compute the training and test error. How do the result compare with part a.

    **Solution:**

#### Q2

Using the training data `my_sahp_train` to fit LDA and QDA models of `expensive` on all three variables (`gar_car`, `liv_area`, `oa_qual`). Compute the training and test error. How do the results compare with Q1?

**Solution:**

#### Q3

Q6 in Chapter 4 of ISLRv2 (Page 191). Hint: Use equation (4.7) on Page 137 of ISLRv2.

**Solution:**
